---
title: "Reproducible Research"
author: "Michael Kroog"
date: "July 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reproducible Research Week 2 Project 1
==========================================================

Calculate the total number of steps taken per day

```{r}
activity <- read.csv("activity.csv", header = TRUE)
totalsum <- with(activity, tapply(steps, date, sum, na.rm = T))
dfsum <- data.frame(date = names(totalsum), steps = totalsum)
```

Histogram of the total number of steps taken per day

```{r}
library(ggplot2)
ggplot(data = dfsum, aes(x = steps)) + geom_histogram(bins = 30) + xlab("Total Steps")
```

The mean and median of the total number of steps taken per day

```{r}
totalmean <- with(activity, tapply(steps, date, mean, na.rm = T))
dfmean <- data.frame(date = names(totalmean), steps = totalmean)
dfmean <- na.omit(dfmean)

totalmedian <- with(activity, tapply(steps, date, median, na.rm = T))
dfmedian <- data.frame(date = names(totalmedian), steps = totalmedian)
dfmedian <- na.omit(dfmedian)
```
The mean number of steps taken per day

```{r}
dfmean
```

The median number of steps taken per day

```{r}
dfmedian
```

Time series plot of the interval and average number of steps taken per day

```{r}
totalinterval <- with(activity, tapply(steps, interval, mean, na.rm = T))
dfinterval <- data.frame(steps = totalinterval, interval = names(totalinterval))

ggplot(dfinterval, aes(x = interval, y = steps)) + geom_line(aes(group = 1), color = "blue") + scale_x_discrete(breaks = seq(60, 2355, 60), limits = 0:2355) + theme(axis.text.x= element_text(size = 10, angle = 90)) + xlab("Interval") + ylab("Average Number of Steps")
```

The plot shows that interval 835 contains on average the maximum number of steps (206)

Total number of missing values

```{r}
sum(is.na(activity))
```
The total number of missing vales is 2304

The following code finds the rows containing NA values, creates a data frame
of equal length containing the average number of steps taken for the corressponding
interval of missing values and imputes them in place of the NAs

```{r}
activitynew <- read.csv("activity.csv", header = TRUE)
miss <- which(is.na(activitynew$steps))
dfintervalrep <- sapply(dfinterval, rep.int, times = 4)
activitynew$steps[miss] <- dfintervalrep

activitymean <- with(activitynew, tapply(steps, date, mean, na.rm = T))
dfactivitymean <- data.frame(date = names(activitymean), steps = activitymean)

activitymedian <- with(activitynew, tapply(steps, date, median, na.rm = T))
dfactivitymedian <- data.frame(date = names(activitymedian), steps = activitymedian)

ggplot(data = dfactivitymean, aes(x = steps)) + geom_histogram(bins = 30) + xlab("Total Steps")
```

Imputed mean

```{r}
dfactivitymean
```

Imputed median

```{r}
dfactivitymedian

identical(dfmean, dfactivitymean)
identical(dfmedian ,dfactivitymedian)
```

The values of the means and medians of original data are different from the values of the adjusted data.

```{r}
values <- rep("missing", times = 53)
values <- append(values, rep("imputed", times =61))
newdfmean <- rbind(dfmean, dfactivitymean)
newdfmean$values <- factor(cbind(values))
```

The impact of adding values in place of NAs results are displayed in the following plot

```{r}
ggplot(newdfmean, aes(x = date, y = steps, color = values)) + geom_line(aes(group = 1)) + facet_grid(values ~ .) + xlab("Date") + ylab("Average Number of Steps") + theme(axis.text.x = element_text(size = 10, angle = 90)) + scale_x_discrete(expand=c(0.001, 0))
```

Creates a new factor variable indicating whether a day is a weekday or weekend
```{r}
activityweek <- activitynew
activityweek$weekday <- weekdays(as.Date(activityweek$date))
activityweek$weekday <- gsub("S.*y","Weekend", activityweek$weekday)
activityweek$weekday <- factor(gsub("[MTWF].*y","Weekday", activityweek$weekday))
```

Time series panel plot of 5-minute intervals and average steps taken

```{r}
activityweekend <- activityweek[which(activityweek$weekday %in% "Weekend"),]
activityweekday <- activityweek[which(activityweek$weekday %in% "Weekday"),]

activityweekendmean <- with(activityweekend, tapply(steps, interval, mean))
activityweekendmean <- data.frame(steps = activityweekendmean, interval = names(activityweekendmean))
activityweekdaymean <- with(activityweekday, tapply(steps, interval, mean))
activityweekdaymean <- data.frame(steps = activityweekdaymean, interval = names(activityweekdaymean))

activityweekendmean$weekday <- cbind(rep("Weekend", each = 288))
activityweekdaymean$weekday <- cbind(rep("Weekday", each = 288))

dfweekmean <- rbind(activityweekendmean, activityweekdaymean)
dfweekmean$weekday <- factor(dfweekmean$weekday)

ggplot(dfweekmean, aes(x = interval, y = steps, color = weekday)) + geom_line(aes(group = 1)) + scale_x_discrete(breaks = seq(0, 2355, 60), limits = 0:2355) + theme(axis.text.x= element_text(size = 10, angle = 90)) + facet_grid(weekday ~ .) + xlab("Interval") + ylab("Average Number of Steps")
```